---
title: "01_SimpleUseCase"
output: rmarkdown::github_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "rendered") })
---

# Simple RStudio Use Case

Clear the workspace
```{r}
rm(list = ls())
```

```{r, message=FALSE}
if (!'tidymodels' %in% installed.packages()) {install.packages('tidymodels', repos='https://packagemanager.rstudio.com/all/__linux__/bionic/latest')}
```

```{r, message=FALSE}
if (!'tidyverse' %in% installed.packages()) {install.packages('tidyverse')}
suppressWarnings(library(tidyverse))
suppressWarnings(library(tidymodels))
```


## Downloading and Processing the Dataset
The model uses the [abalone dataset](https://archive.ics.uci.edu/ml/datasets/abalone)
from the UCI Machine Learning Repository.
```{r, message=FALSE}
abalone <- read_csv(file = 'dataset/abalone.csv', col_names = FALSE)
names(abalone) <- c('sex', 'length', 'diameter', 'height', 'whole_weight', 'shucked_weight', 'viscera_weight', 'shell_weight', 'rings')
head(abalone)
```

The output above shows that `sex` is a factor data type but is currently a
character data type (F is Female, M is male, and I is infant). Change `sex` to a
factor and view the statistical summary of the dataset:
```{r}
abalone <- abalone %>%
  mutate(sex = as_factor(sex))
```

```{r}
summary(abalone)
```

The summary above shows that the minimum value for `height` is 0.

Visually explore which abalones have `height` equal to 0 by plotting the
relationship between `rings` and `height` for each value of `sex`:
```{r}
abalone %>%
  ggplot(aes(x = height, y = rings, color = sex)) +
  geom_point() +
  geom_jitter() +
  theme_minimal()
```

The plot shows multiple outliers: two infant abalones with a height of 0 and a
few female and male abalones with greater heights than the rest. Let’s filter
out the two infant abalones with a height of 0.

```{r}
abalone <- abalone %>%
  filter(height != 0)

head(abalone)
```


## Preparing the Dataset for Model Training

The model needs three datasets: one for training, testing, and validation.
First, convert `sex` into a dummy variable and move the target, `rings`, to the
first column. Amazon SageMaker algorithm require the target to be in the first
column of the dataset.
```{r}
abalone <- abalone %>%
  mutate(female = as.integer(ifelse(sex == 'F', 1, 0)),
         male = as.integer(ifelse(sex == 'M', 1, 0)),
         infant = as.integer(ifelse(sex == 'I', 1, 0))) %>%
  select(-sex)
abalone <- abalone %>%
  select(rings:infant, length:shell_weight)
head(abalone)
```

Sample data for model training
```{r}
set.seed(42)
abalone_train <- abalone %>%
  sample_frac(size = 0.7)
abalone <- anti_join(abalone, abalone_train)
abalone_test <- abalone %>%
  sample_frac(size = 0.5)
abalone_valid <- anti_join(abalone, abalone_test)
```

```{r}

lm_model <- linear_reg() %>%
            set_engine('lm') %>% 
            set_mode('regression')
```

```{r}

lm_fit <- lm_model %>%
          fit(rings ~ ., data = abalone_train)
```

```{r}
summary(lm_fit$fit)
```


```{r}
test_results <- predict(lm_fit, new_data = abalone_test) %>%
                            bind_cols(abalone_test)
head(test_results)
```

```{r}
ggplot(data = test_results,
       mapping = aes(x = .pred, y = rings)) +
  geom_point(color = '#006EA1') +
  geom_abline(intercept = 0, slope = 1, color = 'orange') +
  labs(title = 'Linear Regression Results',
       x = 'Predicted Rings',
       y = 'Actual Rings')
```

